Privacy-Respecting Type Error Telemetry at Scale

### 2023-10-01 programming23/paper.tex
reorder cites

% TODO BEFORE SUBMISSION
% \usepackage[sort]{natbib} or otherwise manually fix cites to be in ascending order

@@

Beyond their immediate revelance to {Luau},
answers to these questions have broad implications for the design
of gradually-typed languages.
Luau represents a large-scale combination of ideas from
gradual typing~\cite{st-sfp-2006,tfffgksst-snapl-2017,bat-ecoop-2014},
success typing~\cite{lindahl2006practical},
and semantic subtyping~\cite{CF05:GentleIntroduction,Jef22:SemanticSubtyping}.
Lessons from this experience can inform future applications.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{m-popl-1990,ktu-caap-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.

@@

One way to futher strengthen the privacy of any telemetry design, including
transparent telemetry, is to systematically add noise to the data using
techniques from differential
privacy~\cite{zhlbr-cc-2020,epk-ccs-2014,wblj-usenix-2017}.
Recent work about session-based data that could be added
to Roblox Studio shows how to add noise to traces~\cite{zhlbr-cc-2020},
how to account for known relationships between events~\cite{zhlbr-oopsla-2020},
and how to choose the noise on the client side~\cite{hlzbr-ecoop-2021}.
Applying differential privacy is subtle, however.
While we could fuzz the total number of type errors, if we were to
fuzz the number of \code{CodeTooComplex} errors, it might drastically
affect our inferences.

@@

Our approach to data analysis is inspired by Marceau
et~al.~\cite{mfk-onward-2011,mfk-sigcse-2011} and
Macedo et~al.~\cite{mcpcsprs-abz-2020,mcpcsprs-scp-2021}
for Racket and Alloy4Fun, respectively. Other more recent work has
also studied edit sequences to infer intent~\cite{wk-koli-2020,lgk-pj-2022,rsgl-cpp-2020}.
A major difference is that we have only samples, not full sequences,
which significantly impacts our study methods and inferences.

@@

Much work~\cite{empirical-types,heeren-thesis} on the benefits of static types focuses on
small case studies~\cite{w-popl-1986,hw-scp-2004,td-tosem-2001},
inverviews~\cite{cdhhjklwya-hatra-2020,gstf-hatra-2021,cams-oopsla-2020},
or static
corpuses~\cite{rppd-fse-2014,bhmvv-toplas-2019}
Controlled experiments are rare;
Kleinschmager, et~al.\cite{khrts-icpc-2012} is a notable exception.
Lerner et~al.~\cite{lfgc-pldi-2007} apply a type error repair tool on
over two thousand programs.
Seidel et~al.~\cite{sscwj-oopsla-2017,sjw-jfp-2018} use five thousand programs
to train data-driven methods for localizing errors.
These are all on a vastly smaller scale than our study and also study
different issues.

>>>>

% TODO BEFORE SUBMISSION

@@

Beyond their immediate revelance to {Luau},
answers to these questions have broad implications for the design
of gradually-typed languages.
Luau represents a large-scale combination of ideas from
gradual typing~\cite{bat-ecoop-2014,st-sfp-2006,tfffgksst-snapl-2017},
success typing~\cite{lindahl2006practical},
and semantic subtyping~\cite{CF05:GentleIntroduction,Jef22:SemanticSubtyping}.
Lessons from this experience can inform future applications.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{ktu-caap-1990,m-popl-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.

@@

One way to futher strengthen the privacy of any telemetry design, including
transparent telemetry, is to systematically add noise to the data using
techniques from differential
privacy~\cite{epk-ccs-2014,wblj-usenix-2017,zhlbr-cc-2020}.
Recent work about session-based data that could be added
to Roblox Studio shows how to add noise to traces~\cite{zhlbr-cc-2020},
how to account for known relationships between events~\cite{zhlbr-oopsla-2020},
and how to choose the noise on the client side~\cite{hlzbr-ecoop-2021}.
Applying differential privacy is subtle, however.
While we could fuzz the total number of type errors, if we were to
fuzz the number of \code{CodeTooComplex} errors, it might drastically
affect our inferences.

@@

Our approach to data analysis is inspired by Marceau
et~al.~\cite{mfk-sigcse-2011,mfk-onward-2011} and
Macedo et~al.~\cite{mcpcsprs-abz-2020,mcpcsprs-scp-2021}
for Racket and Alloy4Fun, respectively. Other more recent work has
also studied edit sequences to infer intent~\cite{lgk-pj-2022,rsgl-cpp-2020,wk-koli-2020}.
A major difference is that we have only samples, not full sequences,
which significantly impacts our study methods and inferences.

@@

Much work~\cite{heeren-thesis,empirical-types} on the benefits of static types focuses on
small case studies~\cite{hw-scp-2004,td-tosem-2001,w-popl-1986},
inverviews~\cite{cdhhjklwya-hatra-2020,cams-oopsla-2020,gstf-hatra-2021},
or static
corpuses~\cite{bhmvv-toplas-2019,rppd-fse-2014}
Controlled experiments are rare;
Kleinschmager, et~al.\cite{khrts-icpc-2012} is a notable exception.
Lerner et~al.~\cite{lfgc-pldi-2007} apply a type error repair tool on
over two thousand programs.
Seidel et~al.~\cite{sjw-jfp-2018,sscwj-oopsla-2017} use five thousand programs
to train data-driven methods for localizing errors.
These are all on a vastly smaller scale than our study and also study
different issues.
